Demes: a specification for interchanging complex demography

### 2021-06-27 paper.tex
incorporating comments, fixing typos

In Demes, demographic models consist of one or more interacting populations (or
``demes"; to avoid confusion with the name of the specification itself we will
use the term ``population'' in this discussion, in the understanding that the
terms are interchangeable for our purposes). Populations are defined as a
collection of exchangable individuals that exist a specified period of time,
following a well-defined set of rules regarding population sizes, mating
systems, etc. Ancestor/descendent relationships between populations can be
defined, as well as continuous or instantaneous migration between coexisting
populations.

Population and event times are written as units in the past, so that time zero
corresponds to the final generation or ``now'', and event times in the past are
values greater than zero with larger values for events that occur in the more
distant past. While a natural specification for time units is in generations,
other time units are allowed, such as years. In the case the the time units are
not given in generations, the generation time must also be specified so that
times can be converted into generations. In general, as time flows from the
past to the present, populations, epochs, and migration events should be
specified in their order of appearance, so that their times are in descending
order.

@@

Within populations, population sizes and mating systems are specified within
epochs. Epochs are non-overlapping intervals that cover the entire existence
interval of the population. Each epoch specifies the population size over that
interval, which can be a constant value or function defined by start and end
sizes that must remain positive.  If an epoch has a start time of infinity, the
population size for that epoch must be constant. Epochs can also specify
parameters for nonrandom mating, such as selfing or cloning rates, which give
the probability that offspring are generated from one generation to the next by
self-fertilization or cloning of an indivual \textbf{need to precisely specify
once this is determined}.

Finally, individuals in a population may have parents from a different
population through migrations. These can be defined as continuous migration
rates over intervals that populations overlap in time or through instantaneous
(or pulse) migration events at a given time. Continuous migration rates are
defined as the probability that parents in the ``destination`` population are
chosen from the ``source`` population.  Migration rates are thus per-generation
and must be less than one. Furthermore, if more than one source population have
continuous migration into the same destination population, the sum of those
migration rates must also be less than one, as rates define probabilities. The
probability that parents come from the same population is just one minus the
some of incoming migration rates. On the other hand, pulse migration events
specify the instantaeous replacement of a given fraction of individuals in a
destination population by individuals with parents from a source population.
The fraction must be between zero and one, and if more than one source
population has a pulse migration event to the same destination population and
the same time, those replacement events are applied sequentially in the order
that they are specified in the model.

>>>>

In Demes, demographic models consist of one or more interacting populations (or
``demes"; to avoid confusion with the name of the specification itself we will
use the term ``population'' in this discussion, in the understanding that the
terms are interchangeable for our purposes). Populations are defined as a
collection of exchangable individuals that exist for a specified period of
time, following a well-defined set of rules regarding population sizes, mating
systems, etc. Ancestor/descendant relationships between populations can be
defined, as well as continuous or instantaneous migration between coexisting
populations.

Population and event times are written as units in the past, so that time zero
corresponds to the final generation or ``now'', and event times in the past are
values greater than zero with larger values for events that occur in the more
distant past. While a natural specification for time units is in generations,
other time units are allowed, such as years. In the case that the time units
are not given in generations, the generation time must also be specified so
that times can be converted into generations. In general, as time flows from
the past to the present, populations, epochs, and migration events should be
specified in their order of appearance, so that their times are in descending
order.

@@

Within populations, population sizes and mating systems are specified within
epochs. Epochs are non-overlapping intervals of time that cover the entire
existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.  If an epoch has a start time of
infinity, the population size for that epoch must be constant. Epochs can also
specify parameters for nonrandom mating, such as selfing or cloning rates,
which give the probability that offspring are generated from one generation to
the next by self-fertilization or cloning of an individual \textbf{need to
precisely specify once this is determined}.

Finally, individuals in a population may have parents from a different
population through migrations. These can be defined as continuous migration
rates over time intervals for which populations coexist or through
instantaneous (or pulse) migration events at a given time. Continuous migration
rates are defined as the probability that parents in the ``destination''
population are chosen from the ``source'' population.  Migration rates are thus
per generation and must be less than one. Furthermore, if more than one source
population have continuous migration into the same destination population, the
sum of those migration rates must also be less than one, as rates define
probabilities. The probability that parents come from the same population is
just one minus the sum of incoming migration rates. On the other hand, pulse
migration events specify the instantaeous replacement of a given fraction of
individuals in a destination population by individuals with parents from a
source population.  The fraction must be between zero and one, and if more than
one pulse occurs at the same time, those replacement events are applied
sequentially in the order that they are specified in the model.
